---
title: "midcourse"
output: html_document
date: '2023-01-04'
---
#Load Libraries
```{r}
library(tidyverse)
library(dplyr)
library(data.table)
library(purrr)
library(sf)
library(leaflet)
```


#Load Field of study data
```{r}
na_str <- c("PrivacySuppressed","NULL")
colsfos <- c(1:4, 6:14, 42:44, 72, 74,76:101, 110:113)

fos16<- fread('data/FieldOfStudyData1415_1516_PP.csv', na = na_str, select = colsfos)%>% 
  mutate(org_yr = "1415_1516", year = rep(c(2016)))
  

fos17<- fread('data/FieldOfStudyData1516_1617_PP.csv', na = na_str, select = colsfos)%>% 
  mutate(org_yr = "1516_1617", year = rep(c(2017)))


fos18<- fread('data/FieldOfStudyData1617_1718_PP.csv', na = na_str, select = colsfos)%>% 
  mutate(org_yr = "1617_1718", year = rep(c(2018)))


fos19<- fread('data/FieldOfStudyData1718_1819_PP.csv', na = na_str, select = colsfos)%>% 
  mutate(org_yr = "1718_1819", year = rep(c(2019)))
```


#List of column names
```{r}
colnames(fos16)
```


#join dataframes
```{r}
fos_all <- rbind(fos16, fos17, fos18, fos19)
```

#Filter out foreign universities
```{r}
fos_all %>% group_by(CONTROL) %>% summarise(n_distinct(INSTNM))
```
```{r}
fos_all <- fos_all %>% filter(CONTROL != "Foreign")
```

#Get dataframe information
```{r}
str(fos_all)
```
#Count how many na per column
```{r}
data.frame(sapply(fos_all, function(y) sum(length(which(is.na(y)))))) %>% view()
```

```{r}
fos_all %>% group_by(CREDDESC, CIPDESC) %>% summarise(nu_awards = sum(IPEDSCOUNT2, na.rm = T))
```




```{r}
fos_all %>% 
  ggplot(aes(x = year, y= EARN_MDN_HI_1YR))+
  geom_bar(stat = "identity")
```
#comparative barchart for 1yr, 2yr, 3yr earnings
```{r}
fos_all %>% 
  ggplot(aes(x = org_yr, y= EARN_MDN_HI_2YR))+
  geom_bar(stat = "identity")
```





#Read merged(instituition level data)
```{r}
colstokeep <- c(1:8, 11:12, 15:17, 19:23,37:38, 45:47, 56:59, 61:99, 290:388, 397:426, 429:432,
                438:439, 446,449,459, 462,  472, 475, 485, 488,
                524, 527, 537, 540, 550, 553, 563, 566, 576, 579, 589, 592,
                602, 605, 615, 618, 1408:1427, 1432:1447, 1462:1475,
                1504:1531, #median and n debt breakdown
                1605:1607, 1609:1617, #demographics
                1627:1629, #census unemp, povrate, hhincome 
                1742:1744, 1809:1812, 
                2042:2044, #number employed 3yrs after comp
                2394:2399, #all
                2955:2989 #earnings by yr since completion, income
                )
m97 <- fread("data/MERGED1996_97_PP.csv", na = na_str, select = colstokeep) %>%mutate(year = rep(c(1997))) 
m98 <- fread("data/MERGED1997_98_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(1998)))
m99 <- fread("data/MERGED1998_99_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(1999)))
m00 <- fread("data/MERGED1999_00_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(2000)))

m01 <- fread("data/MERGED2000_01_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(2001)))
m02 <- fread("data/MERGED2001_02_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(2002)))
m03 <- fread("data/MERGED2002_03_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(2003)))
m04 <- fread("data/MERGED2003_04_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(2004)))
m05 <- fread("data/MERGED2004_05_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(2005)))
m06 <- fread("data/MERGED2005_06_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(2006)))
m07 <- fread("data/MERGED2006_07_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(2007)))
m08 <- fread("data/MERGED2007_08_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(2008)))
m09 <- fread("data/MERGED2008_09_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(2009)))
m10 <- fread("data/MERGED2009_10_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(2010)))

m11 <- fread("data/MERGED2010_11_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(2011)))
m12 <- fread("data/MERGED2011_12_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(2012)))
m13 <- fread("data/MERGED2012_13_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(2013)))
m14 <- fread("data/MERGED2013_14_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(2014)))
m15 <- fread("data/MERGED2014_15_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(2015)))
m16 <- fread("data/MERGED2015_16_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(2016)))
m17 <- fread("data/MERGED2016_17_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(2017)))
m18 <- fread("data/MERGED2017_18_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(2018)))
m19 <- fread("data/MERGED2018_19_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(2019)))
m20 <- fread("data/MERGED2019_20_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(2020)))
m21 <- fread("data/MERGED2020_21_PP.csv", na = na_str, select = colstokeep)%>% mutate(year = rep(c(2021)))

```


```{r}
data.frame(sapply(m97, function(y) sum(length(which(is.na(y)))))) %>% view()
```

```{r}
df_all <- mget(ls(pattern = "^m\\d{2}"))
```

```{r}
new_list <- lapply(df_all, function(x) filter(x, ((PREDDEG>1)&(HIGHDEG>1)&(PREDDEG<4))))

list2env(new_list, environment())
```



#Row bind Inst Data

```{r}
inst_data <- do.call(rbind, new_list)
```
```{r}
sapply(inst_data, summary)
```

```{r}
data.frame(sapply(inst_data, function(y) sum(length(which(is.na(y)))))) %>% view()
```

#Dictionary/Key for column names
```{r}
dictcsv<- read_csv("data/dictionary.csv")
dictcsv<- dictcsv %>% drop_na(`VARIABLE NAME...2`)
```


#Change column names
```{r}
setnames(inst_data, old = dictcsv$`VARIABLE NAME...2`, new = dictcsv$`developer-friendly name...6`, skip_absent = TRUE)
```




#Check what program year institutions are
```{r}
inst_data %>% filter(!is.na(avg_net_price.program_year)) %>% select(name, avg_net_price.program_year, predomdegawrd, highestdegawrd) %>% view()
```

#Mutate new columns based on key
```{r}
inst_data <- inst_data %>% select(-degree_urbanization) %>% 
  mutate(highestdegawrd = case_when(
      degrees_awarded.highest == 2 ~ "Associate",
      degrees_awarded.highest == 3 ~ "Bachelors",
      degrees_awarded.highest == 4 ~ "Graduate"),
    predomdegawrd = case_when(
      degrees_awarded.predominant == 2 ~ "Predominantly Associate",
      degrees_awarded.predominant == 3 ~ "Predominantly Bachelors",
      degrees_awarded.predominant == 4 ~ "Entirely Graduate"),
    public_private = case_when(
      ownership == 1 ~ "Public",
      ownership == 2 ~ "Private nonprofit",
      ownership == 3 ~ "Private for-profit"),
    region = case_when(
      region_id == 0 ~ "US Service Schools", 
      region_id == 1 ~ "New England",
      region_id == 2 ~ "Mid East",
      region_id == 3 ~ "Great Lakes",
      region_id == 4 ~ "Plains",
      region_id == 5 ~ "Southeast",
      region_id == 6 ~ "Southwest",
      region_id == 7 ~ "Rocky Mountains",
      region_id == 8 ~ "Far West",
      region_id == 9 ~ "Outlying Areas")
    
    )
```




```{r}
inst_data %>% group_by(locale) %>% summarise(n_distinct(id))
```
#export list of column names and how many nas

```{r}
# columnlist <- data.frame(sapply(inst_data, function(y) sum(length(which(is.na(y))))))
# columnlist<- cbind("name" =rownames(columnlist), columnlist)
# 
# library(writexl)
# writexl::write_xlsx(columnlist, "columnlist.xlsx")
```




#Data Exploration

```{r}
inst_data %>% 
  group_by(year) %>% 
  summarise(publicv = sum(title_iv.public.all, na.rm = T)) %>% 
  ggplot(aes(x=year, y=publicv))+
  geom_line()
  
```
```{r}
inst_data %>% group_by(year) %>% summarize(sum(title_iv.public.all, na.rm = T))
```
#Time Series by Public and Private TitleIV
```{r}
inst_data %>% 
  group_by(year) %>% 
  summarise(publicv = sum(title_iv.public.all, na.rm = T),
            private = sum(title_iv.private.all, na.rm = T)) %>% 
  ggplot(aes(x=year))+
  geom_line(aes(y=publicv), color = "red")+
  geom_point(aes(y=publicv), color = "red")+
  geom_line(aes(y=private), color = "blue")+
  geom_point(aes(y=private), color = "blue")+
  scale_x_continuous(breaks = seq(1995, 2022, 2))+
  ylab("enrollment")+
  ggtitle("Title IV School Enrollments Private vs Public")
  
```
```{r}
inst_data %>% 
  group_by(year) %>% 
  summarise(enrollmentsum = mean(enrollment.all, na.rm = T)) %>% 
  ggplot(aes(x=year))+
  geom_line(aes(y=enrollmentsum), color = "red")+
  geom_point(aes(y=enrollmentsum))
  
  
```
```{r}
inst_data %>% 
  group_by(year) %>% 
  summarise(instatetuit = mean(tuition.in_state, na.rm = T),
            instexpend = mean(instructional_expenditure_per_fte, na.rm = T)) %>% 
  ggplot(aes(x=year))+
  geom_line(aes(y=instatetuit), color = "red")+
  geom_point(aes(y=instatetuit))+
  geom_line(aes(y=instexpend))
```

```{r}
inst_data %>% arrange(desc(instructional_expenditure_per_fte)) %>% select(name, instructional_expenditure_per_fte) %>% view()
```

#### Price/tuition

```{r}
inst_data %>% 
  group_by(year) %>% 
  summarise(instatetuit = median(tuition.in_state, na.rm = T),
            instexpend = median(instructional_expenditure_per_fte, na.rm = T),
            outstatetuit = median(tuition.out_of_state, na.rm = T),
            costattendance = median(attendance.academic_year, na.rm = T)) %>% 
  ggplot(aes(x=year))+
  geom_line(aes(y=instatetuit), color = "red")+
  geom_point(aes(y=instatetuit))+
  geom_line(aes(y=instexpend))+
  geom_line(aes(y=outstatetuit), color = "blue")+
  geom_point(aes(y=outstatetuit))+
  geom_line(aes(y=costattendance), color = "green")+
  geom_point(aes(y=costattendance))+
  ggtitle("Instate tuition and instructional expenditure per fte")
```


```{r}
inst_data %>% filter(!is.na(tuition.program_year))
```

```{r}
inst_data %>%group_by(year) %>%
  summarise(li = mean(share_lowincome.0_30000, na.rm = T),
            mi = mean(share_middleincome.30001_48000, na.rm = T),
            mi2= mean(share_middleincome.48001_75000, na.rm = T),
            hi = mean(share_highincome.75001_110000, na.rm = T),
            hi2 = mean(share_highincome.110001plus, na.rm = T)) %>% 
  ggplot(aes(x=year))+
  geom_line(aes(y=li), color = "red")+
  geom_line(aes(y=mi), color = "blue")+
  geom_line(aes(y=mi2), color = "darkgreen")+
  geom_line(aes(y=hi), color = "purple")+
  geom_line(aes(y=hi2), color = "orange")
```


```{r}
inst_data %>%group_by(year) %>%
  summarise(li = mean(share_dependent_lowincome.0_300000, na.rm = T),
            mi = mean(share_dependent_middleincome.300001_48000, na.rm = T),
            mi2= mean(share_dependent_middleincome.48001_75000, na.rm = T),
            hi = mean(share_dependent_highincome.75001_110000, na.rm = T),
            hi2 = mean(share_dependent_highincome.110001plus, na.rm = T)) %>% 
  ggplot(aes(x=year))+
  geom_line(aes(y=li), color = "red")+
  geom_line(aes(y=mi), color = "blue")+
  geom_line(aes(y=mi2), color = "darkgreen")+
  geom_line(aes(y=hi), color = "purple")+
  geom_line(aes(y=hi2), color = "orange")
```


```{r}
inst_data %>%group_by(year) %>%
  summarise(li = mean(share_independent_lowincome.0_30000, na.rm = T),
            mi = mean(share_independent_middleincome.30001_48000, na.rm = T),
            mi2= mean(share_independent_middleincome.48001_75000, na.rm = T),
            hi = mean(share_independent_highincome.75001_110000, na.rm = T),
            hi2 = mean(share_independent_highincome.110001plus, na.rm = T)) %>% 
  ggplot(aes(x=year))+
  geom_line(aes(y=li), color = "red")+
  geom_line(aes(y=mi), color = "blue")+
  geom_line(aes(y=mi2), color = "darkgreen")+
  geom_line(aes(y=hi), color = "purple")+
  geom_line(aes(y=hi2), color = "orange")
```

```{r}
inst_data %>%group_by(year) %>%
  summarise(wht = mean(drewht, na.rm = T),
            blk = mean(demographics.race_ethnicity.black, na.rm = T),
            his= mean(demographics.race_ethnicity.hispanic, na.rm = T),
            asi = mean(demographics.race_ethnicity.asian, na.rm = T)) %>% 
  ggplot(aes(x=year))+
  geom_line(aes(y=wht), color = "red")+

  geom_line(aes(y=blk), color = "blue")+
  
  geom_line(aes(y=his), color = "purple")+
  geom_line(aes(y=asi), color = "orange")+
  scale_x_continuous(breaks = seq(2000, 2022, 2))
```
```{r}
inst_data <- inst_data %>% mutate(drewht=if_else(is.na(demographics.race_ethnicity.white),
                               demographics.race_ethnicity.white_2000, demographics.race_ethnicity.white)) 
```

```{r}
inst_data %>% select(year, drewht, demographics.race_ethnicity.white_2000, demographics.race_ethnicity.white) %>% arrange(demographics.race_ethnicity.white_2000) %>% view()
```


```{r}
inst_data %>% mutate(drewht=if_else(is.na(demographics.race_ethnicity.white),
                               demographics.race_ethnicity.white_2000, demographics.race_ethnicity.white)) 
```

```{r}
inst_data %>%group_by(year) %>%
  summarise(avg = mean(share_firstgeneration, na.rm = T),
            mi = mean(share_firstgeneration_parents.middleschool, na.rm = T),
            hi= mean(share_firstgeneration_parents.highschool, na.rm = T),
            col = mean(share_firstgeneration_parents.somecollege, na.rm = T)) %>% 
  ggplot(aes(x=year))+
  geom_line(aes(y=avg), color = "red")+
  geom_line(aes(y=mi), color = "blue")+
  
  geom_line(aes(y=hi), color = "purple")+
  geom_line(aes(y=col), color = "orange")+
  ggtitle("Enrollment by First Generation")
```

```{r}
inst_data %>%group_by(year) %>%
  summarise(avg = mean(title_iv.withdrawn_by.2yrs, na.rm = T),
            mi = mean(title_iv.still_enrolled_by.2yrs, na.rm = T)) %>% 
  ggplot(aes(x=year))+
  geom_line(aes(y=avg), color = "red")+
  geom_point(aes(y=avg), color = "red")+
  geom_line(aes(y=mi), color = "blue")+
  geom_point(aes(y=mi), color = "blue")
  
  
```


```{r}
inst_data %>%group_by(year) %>%
  summarise(liwd = mean(title_iv.low_inc.withdrawn_by.2yrs, na.rm = T),
            lien = mean(title_iv.low_inc.still_enrolled_by.2yrs, na.rm = T),
            midwd = mean(title_iv.mid_inc.withdrawn_by.2yrs, na.rm = T),
            midenr = mean(title_iv.mid_inc.still_enrolled_by.2yrs, na.rm = T),
            hiwd = mean(title_iv.high_inc.withdrawn_by.2yrs, na.rm = T),
            hien = mean(title_iv.high_inc.still_enrolled_by.2yrs, na.rm = T)
            ) %>% 
  ggplot(aes(x=year))+
  geom_line(aes(y=liwd), color = "purple",linetype = "dashed", size = 1)+
  geom_point(aes(y=liwd), color = "purple", size = 1)+
  geom_line(aes(y=lien), color = "purple", size =0.75)+
  #geom_point(aes(y=mi), color = "royalblue4")+
  geom_line(aes(y=midwd), color = "chartreuse4", linetype = "dashed", size=0.75)+
  geom_line(aes(y=midenr), color = "chartreuse4", size = 0.75)+
  geom_line(aes(y=hiwd), color = "darkgoldenrod1", linetype = "dashed", size=0.75)+
  geom_line(aes(y=hien), color = "darkgoldenrod1", size = 0.75)+
  scale_x_continuous(breaks=seq(2000, 2022, 1))+
  theme_minimal()
  
  
```

```{r}
inst_data %>%group_by(year) %>%
  summarise(fewd = mean(title_iv.female.withdrawn_by.2yrs, na.rm = T),
            feen = mean(title_iv.female.still_enrolled_by.2yrs, na.rm = T),
            mawd = mean(title_iv.male.withdrawn_by.2yrs, na.rm = T),
            maen = mean(title_iv.male.still_enrolled_by.2yrs, na.rm = T),
            ) %>% 
  ggplot(aes(x=year))+
  geom_line(aes(y=fewd), color = "purple" , linetype = "dashed", size = 1)+
  
  geom_line(aes(y=feen), color = "purple", size =0.75)+
  #geom_point(aes(y=mi), color = "royalblue4")+
  geom_line(aes(y=mawd), color = "chartreuse4", linetype = "dashed", size=0.75)+
  geom_line(aes(y=maen), color = "chartreuse4", size = 0.75)+
  
  scale_x_continuous(breaks=seq(2000, 2022, 1))+
  theme_minimal()
  
```

```{r}
#write_csv(inst_data,"inst_data.csv")
```

```{r}
latlngdf <- m21 %>% select(UNITID,ACCREDAGENCY, LATITUDE, LONGITUDE)
```
```{r}
latlngdf <- latlngdf %>% rename("id" = "UNITID")
```

```{r}
inst_data2 <- left_join(inst_data, latlngdf, by="id")
```

```{r}
inst_data2 %>% select(id, operating, name, accreditor, ACCREDAGENCY, location.lat, LATITUDE, location.lon, LONGITUDE ) %>% view
```

```{r}
inst_data2 %>% filter(is.na(ACCREDAGENCY)) %>% select(name, ACCREDAGENCY, LATITUDE, LONGITUDE)%>% view()
```
```{r}
inst_data2<- inst_data2 %>% 
  filter(!is.na(ACCREDAGENCY)) %>% 
  filter(!is.na(LATITUDE)) %>% 
  filter(!is.na(LONGITUDE)) 
```

```{r}
inst_data2sf <- st_as_sf(inst_data2, coords = c("LATITUDE", "LONGITUDE"), crs = 4326)
```


```{r}
write_csv(inst_data2sf,"inst_data2.csv")
```


```{r}
inst_data2sf() %>% 
    leaflet() %>% 
    addProviderTiles(provider = "CartoDB.PositronNoLabels") %>%
    setView(lng = -86.7816, lat = 36.1627, zoom = 12)
```



