---
title: "inst_data_exp"
output: html_document
date: '2023-01-14'
---

```{r}
library(dplyr)
library(scales)
library(sf)
library(tidyverse)
```

```{r}
insdta <-  read_rds('inst_data3.rds')
insdta <- st_drop_geometry(insdta)
```

#Select columns to keep for shiny app
```{r}
inst_data3 <- insdta %>% select(id, name, city, state, public_private, year, size, degrees_awarded.predominant,  predomdegawrd, degrees_awarded.highest, title_iv.female.withdrawn_by.2yrs, title_iv.female.still_enrolled_by.2yrs, title_iv.male.withdrawn_by.2yrs,title_iv.male.still_enrolled_by.2yrs, year, ownership, region_id, region, latt, longg, geometry, state_fips, state_fips2, enrollment.all)
```

#Make rds file for shiny app
```{r}
write_rds(inst_data3,"inst_data3.rds")
```


```{r}
prgms <- insdta %>% select(id, name, city, state, degrees_awarded.predominant, degrees_awarded.highest, 
                           ownership, region_id, size, enrollment.all,  part_time_share, 370:378 
                           ,starts_with("program_percentage."))
```

```{r}
#write_csv(prgms,"prgms.csv")
```

```{r}
programnames <- prgms %>% select(starts_with("program_percentage.")) %>% colnames()
class(programnames)
```


```{r}
prgms %>%
    dplyr::group_by(year, public_private) %>% 
    #mutate(noq = noquote("program_percentage.history")) %>% 
    dplyr::summarise(val = mean(program_percentage.history, na.rm = T)) %>% 
    ggplot(aes(year, val)) +
    geom_line(color = public_private)+
    scale_y_continuous(labels = function(x) paste0(x*100, "%"))
```

```{r}
prgms %>% group_by(year, public_private) %>% summarise(val = mean(program_percentage.history, na.rm = T)) %>% 
  ggplot(aes(year, val))+
  geom_line(aes(color = public_private, group = public_private))
```
```{r}
prgms["public_private"]
    
```

```{r}
prognm2 <- programnames[-39]
prognm2
```

```{r}
prognm2 <- as.list(prognm2)
```



```{r}
colnames(prgms[21:58])
```



```{r}
insdta <- insdta %>%select(-c(2:3)) 
insdta <- insdta %>% select(-c(6, 8, 14:16))

insdta <- insdta %>% select(-c(starts_with("net_price.program_reporter"), 
                     starts_with("net_price.other_acad"),
                     avg_net_price.program_year, avg_net_price.other_academic_year,
                     contains("other_acad_calendar"),
                     contains("program_reporter"))
                    )

insdta <- insdta %>% select(-c(contains("other_academic_year"),
                     contains("program_year"),
                     title_iv.approval_date))

insdta <- insdta %>% select(-ends_with("pre2010"))
insdta <- insdta %>% select(-c(ends_with("_2000"), ends_with("_2009")))
insdta <- insdta %>% select(-c("enrollment.all", ends_with("pooled")))

```


```{r}
insdta <- insdta %>% select(-operating)
```

```{r}
insdta <- insdta%>% mutate("Predom_Degree_Awarded" = case_when(
  predomdegawrd == "Predominantly Associate" ~ "Associate's Degree",
  predomdegawrd == "Predominantly Bachelors" ~ "Bachelor's Degree"
))
```










```{r}
q<- insdta %>% filter(year >2001) %>% 
  group_by(year) %>% 
  summarize(adm_rt = mean(admission_rate.overall, na.rm =T)) %>% 
  ggplot()+
  geom_line(aes(year, adm_rt),size = 1)+
  ylab("Admission Rate")+
  scale_x_continuous(breaks = seq(2000, 2022, 1))+
  coord_cartesian(ylim = c(0.6, 0.85), expand = TRUE) +
  scale_y_continuous( labels = percent)+
   theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))+
  ggtitle('Overall Admission Rate Nationwide')
q
 dev.off()
ggsave(q, width = 10, height = 6, dpi = 300, filename = "AddNationWide.pdf") 
```

```{r}
p <- insdta %>% 
  filter(year >2002) %>% 
  group_by(year, public_private) %>% 
  summarize(adm_rt = mean(admission_rate.overall, na.rm =T)) %>% 
  ggplot()+
  geom_line(aes(x=year, y=adm_rt, color = public_private), size = 1)+
  xlim(2002, 2022)+
  ylab("Admission Rate")+
  #ylim(0.6,0.8)+
  
  coord_cartesian(ylim = c(0.6, 0.85), expand = TRUE) +
  scale_x_continuous(breaks = seq(2002, 2022, 1))+
  scale_y_continuous( labels = percent)+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1),
        legend.title=element_blank())+
  ggtitle('Admission Rate by Ownership')
 
p
dev.off()
ggsave(p, width = 10, height = 6, dpi = 300, filename = "AddOwn.pdf")  

```
```{r}
r <- insdta %>% 
  filter(year >2002) %>% 
  group_by(year, Predom_Degree_Awarded) %>% 
  summarize(adm_rt = mean(admission_rate.overall, na.rm =T)) %>% 
  ggplot()+
  geom_line(aes(x=year, y=adm_rt, color = Predom_Degree_Awarded), size =1)+
  xlim(2002, 2022)+
  #ylim(0.6,0.8)+
  ylab("Admission Rate")+
  theme_minimal()+
  scale_x_continuous(breaks = seq(2002, 2022, 1))+
   coord_cartesian(ylim = c(0.6, 0.85), expand = TRUE) +
  scale_y_continuous( labels = percent)+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1),
        legend.title=element_blank())+
  ggtitle('Admission Rate by School Type')

r
dev.off()
ggsave(r, width = 10, height = 6, dpi = 300, filename = "AddPRED.pdf")  
```

```{r}
insdta %>%
  filter(year >2001) %>% 
  group_by(year, predomdegawrd) %>% 
  summarize(adm_rt = mean(admission_rate.by_ope_id, na.rm =T)) %>% 
  ggplot()+
  geom_line(aes(x=year, y=adm_rt, color = predomdegawrd), size = 1)+
  xlim(2002, 2022)+
  ylim(0.6,0.8)+
  ylab("Admission Rate")+
  theme_minimal()+
  ylim(0.6,0.8)+
  scale_x_continuous(breaks = seq(2002, 2022, 1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1),
        legend.title=element_blank())+
  ggtitle('Admission Rate by School Type')
```





```{r}
insdta %>% 
  filter(year > 2001) %>% 
  group_by(year, predomdegawrd, public_private) %>% 
  summarize(adm_rt = mean(admission_rate.overall, na.rm =T)) %>% 
  ggplot()+
  geom_line(aes(x=year, y=adm_rt, color = public_private))+
  xlim(2002, 2022)+
  ylim(0.6,0.8)+
  
  scale_x_continuous(breaks = seq(2002, 2022, 1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))+
  ggtitle('Overall admission rate by year, grouped by predominant degree')+
  facet_wrap( ~ predomdegawrd)
```
```{r}
insdta %>%
  filter(year >2001) %>% 
  filter(public_private != "Private for-profit") %>% 
  group_by(year, public_private) %>% 
  summarize(act_scr = mean(act_scores.midpoint.cumulative, na.rm =T)) %>% 
  ggplot()+
  geom_line(aes(x=year, y=act_scr, color=public_private))+
  #xlim(2002, 2022)+
  ylim(20,25)+
  
  scale_x_continuous(breaks = seq(2002, 2022, 1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))+
  ggtitle('act midpoint, grouped by type of school')
```

```{r}
insdta %>%
  filter(year >2001) %>% 
  filter(public_private != "Private for-profit") %>% 
  group_by(year, public_private) %>% 
  summarize(sat_avg = mean(sat_scores.average.by_ope_id, na.rm =T)) %>% 
  ggplot()+
  geom_line(aes(x=year, y=sat_avg, color=public_private))+
  #xlim(2002, 2022)+
  #ylim(20,25)+
  
  scale_x_continuous(breaks = seq(2002, 2022, 1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))+
  ggtitle('act midpoint, grouped by type of school')
```




```{r}
insdta %>% 
  group_by(year, highestdegawrd) %>% 
  summarize(adm_rt = mean(admission_rate.overall, na.rm =T)) %>% 
  ggplot()+
  geom_line(aes(x=year, y=adm_rt, color = highestdegawrd))+
  xlim(2002, 2022)+
  ylim(0.6,0.8)+
  
  scale_x_continuous(breaks = seq(2002, 2022, 1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))+
  ggtitle('admission rate by year, grouped by highest degree awarded by institution')
```



#Completion Rates
```{r}
z<-
insdta %>% 
  filter(region_id != 9, region_id != 0) %>% 
  group_by(year, region, public_private) %>% 
  summarise(complrt_150_percent_4yi = mean(completion_rate_4yr_150nt, na.rm = T)) %>% 
  ggplot()+
  geom_line(aes(x= year, y=complrt_150_percent_4yi, color = region), size = 1)+
  facet_wrap(~public_private)+
  ylab("Completion Rate")+
  ylim(0.2, 1)+
  scale_y_continuous(labels = percent)+
 
  theme_minimal()+
   theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1),
        legend.title=element_blank() )+
  
  ggtitle("4-year Schools")
z
dev.off()
ggsave(z, width = 10, height = 6, dpi = 300, filename = "Comp1.pdf")
```
#Overall, completion looks like an upward trend for private nonprofit and public schools. 
#In general, private non-profit schools have higher completion rates, but public schools
#show higher year-to-year upward trend in completion rates.
#There is not a discernable patternfor private for profit schools.

```{r}
y <- insdta %>%
  filter(region_id != 9, region_id != 0) %>% 
  group_by(year, region, public_private) %>% 
  summarise(complrt_150_pct_l4 = mean(completion_rate_less_than_4yr_150nt, na.rm = T)) %>% 
  ggplot()+
  geom_line(aes(x= year, y=complrt_150_pct_l4, color = region), size = 1)+
  facet_wrap(~public_private)+
  ylab("Completion Rate")+
  ylim(0.2, 1)+
  scale_y_continuous(labels = percent)+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1),
        legend.title=element_blank())+
  ggtitle("Less than 4-year Schools")
y

y
dev.off()
ggsave(y, width = 10, height = 6, dpi = 300, filename = "Comp2.pdf")  

```

#Completion rates for less than 4 year schools- private for profit schools show higher
#trends in completion and public schools show the lowest completion rates.


```{r}
#This is useless
# insdta <- insdta %>% 
#   mutate(compl_rt_4yr_150_wht = if_else(
#                                           (is.na(completion_rate_4yr_150_white)),
#                                           mean(c(completion_rate_4yr_150_white,completion_rate_4yr_150_white_pre2010), na.rm=T),
#                                           completion_rate_4yr_150_white,
#                                           ),
#          compl_rt_4yr_150_blk = if_else(
#                                           ((year <2009)&(is.na(completion_rate_4yr_150_black))),
#                                           completion_rate_4yr_150_black_pre2010,
#                                           completion_rate_4yr_150_black, ),
#          compl_rt_4yr_150_hsp = if_else(
#                                           ((year <2009)&(is.na(completion_rate_4yr_150_hispanic))),
#                                           completion_rate_4yr_150_hispanic_pre2010,
#                                           completion_rate_4yr_150_hispanic),
#          compl_rt_4yr_150_aian = if_else(
#                                           ((year <2009)&(is.na(completion_rate_4yr_150_aian))),
#                                           completion_rate_4yr_150_aian_pre2010,
#                                           completion_rate_4yr_150_aian),
#          compl_rt_4yr_150_asian = if_else(
#                                         ((year <2009)&(is.na(completion_rate_4yr_150_asian))),
#                                           completion_rate_4yr_150_api_pre2010,
#                                           completion_rate_4yr_150_asian),
#          compl_rt_4yr_150_nhpi = if_else(
#                                           ((year <2009)&(is.na(completion_rate_4yr_150_nhpi))),
#                                           completion_rate_4yr_150_api_pre2010,
#                                           completion_rate_4yr_150_nhpi)) 
#  
  
  # %>%select(-c(completion_rate_4yr_150_white, completion_rate_4yr_150_white_pre2010,
  #           completion_rate_4yr_150_black,completion_rate_4yr_150_black_pre2010,
  #           completion_rate_4yr_150_hispanic_pre2010, completion_rate_4yr_150_hispanic,
  #           completion_rate_4yr_150_aian_pre2010,completion_rate_4yr_150_aian,
  #           completion_rate_4yr_150_api_pre2010, completion_rate_4yr_150_asian,
  #           completion_rate_4yr_150_api_pre2010, completion_rate_4yr_150_nhpi))
            
            
           
```





```{r}
insdta %>% 
  filter(year >2009) %>% 
  group_by(year) %>% 
   
  summarise(compl_wht= mean(completion_rate_4yr_150_white , na.rm = T),
            compl_blk = mean(completion_rate_4yr_150_black, na.rm = T),
            compl_hsp = mean(completion_rate_4yr_150_hispanic, na.rm = T),
            compl_asian = mean(completion_rate_4yr_150_asian, na.rm = T),
            compl_nra = mean(completion_rate_4yr_150_nonresident.alien, na.rm = T),
            compl_ain = mean(completion_rate_4yr_150_aian)) %>% 
  ggplot(aes(x=year))+
  geom_line(aes(y= compl_wht), size = 1, color = "#e60049")+
  geom_line(aes(y= compl_blk), size = 1, color = "#0bb4ff")+
  geom_line(aes(y= compl_hsp), size = 1, color = "#50e991")+
  geom_line(aes(y=compl_asian),size = 1, color = "#b33dc6" )+
  geom_line(aes(y = compl_ain), size = 1, color = "#ef9b20"  )+
  geom_line(aes(y = compl_nra), size = 1, color = "#e6d800"  )+
  #facet_wrap(~public_private)+
  ggtitle("Completion rates for 4yrs by race")

```
```{r}
insdta %>% 
  filter(year > 2004) %>% 
  group_by(year) %>% 
  
  summarise(ret_4yr_ft= mean(retention_rate.four_year.full_time , na.rm = T),
            ret_l4_ft = mean(retention_rate.lt_four_year.full_time, na.rm = T),
            ret_4yr_pt = mean(retention_rate.four_year.part_time, na.rm = T),
            ret_l4yr_pt = mean(retention_rate.lt_four_year.part_time, na.rm = T)) %>% 
            
  ggplot(aes(x=year))+
  geom_line(aes(y= ret_4yr_ft), size = 1, color = "#e60049")+
  geom_line(aes(y= ret_l4_ft), size = 1, color = "#0bb4ff")+
  geom_line(aes(y= ret_4yr_pt), size = 1, color = "#50e991")+
  geom_line(aes(y=ret_l4yr_pt),size = 1, color = "#b33dc6" )+
  
  #facet_wrap(~public_private)+
  ggtitle("Retention Rates by 4yr/less than 4 yr/full time/part time")

```

#Retention rates are higher for 4yr schools and for students who are also full time,
#retention rates are lower for less than 4 yr schools, and when students are part time.

```{r}
t <- insdta %>% 
  filter(region_id != 9, region_id != 0) %>% 
  filter(public_private != "Private for-profit", year>2001) %>% 
  group_by(year, public_private, state) %>% 
  
  summarise(size= sum(size, na.rm = T)) %>% 
            
  ggplot(aes(x=year))+
  geom_line(aes(y= size, color = public_private), size = 1)+
  theme_minimal()+
  theme(axis.text.x = element_blank(),
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank() )+
  xlab("")+
  ylab("")+
  facet_wrap(~state, nrow = 5)+
  guides(fill=guide_legend(title = ""))+
  guides(color=guide_legend(title = ""))+
  ggtitle("Total Enrollment of Undergraduate Degree-seeking Students")
t
dev.off()
ggsave(t, width = 10, height = 6, dpi = 300, filename = "st.pdf")

```





```{r}
insdta %>% 
  group_by(year, public_private) %>% 
  summarise(number_of_schools = n_distinct(id)) %>% 
  ggplot(aes(x=year, y=number_of_schools, fill=public_private)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.5)+
  scale_fill_brewer(palette = "Set1")
```


```{r}
change <- function(df, oldpref, newpref){
  df %>% 
    rename_at(vars(starts_with(oldpref)), funs(sub(oldpref, newpref, .)))
}
```

```{r}
insdta <- change(insdta, '4_yr_', 'Four_yr_')
```


```{r}
insdta %>% group_by(year) %>% 
  summarise(comp_fg = mean(Four_yr_completion.first_generation_students, na.rm =T),
            comp_nfg = mean(Four_yr_completion.no_first_generation_students, na.rm = T)) %>% 
  ggplot(aes(x=year))+
  geom_line(aes(y=comp_fg), color = "darkorange")+
  geom_point(aes(y=comp_fg), color = "darkorange")+
  geom_line(aes(y=comp_nfg), color = "darkorchid")+
  geom_point(aes(y=comp_nfg), color = "darkorchid")
```


```{r}
four_yr_df_b <- insdta %>% select(id, name, highestdegawrd, predomdegawrd, state_fips2, ownership, public_private, region_id, region, year, starts_with("Four_yr")) %>% select(id, name, highestdegawrd, predomdegawrd, state_fips2, ownership, public_private, region_id, region, year, ends_with("income"))
```

```{r}
four_yr_df_b_pvt <- four_yr_df_b %>% 
  pivot_longer(cols = starts_with("Four"),
               names_to = "fouryr_inc_lvl",
               values_to = "fouryr_inc_lvl_size")
```

```{r}
four_yr_df_b_pvt %>% 
  filter(state_fips2 == 23) %>% 
  ggplot(aes(x=fouryr_inc_lvl, y=fouryr_inc_lvl_size, fill=fouryr_inc_lvl))+
  geom_bar(position = "dodge", fun = "mean", stat = "summary")+
  facet_grid(predomdegawrd~public_private)+
  ggtitle("Put in Shiny:4year cohort average number of students by income level")
```
```{r}
#TIME SERIES
four_yr_df_b_pvt %>% 
  #filter(state_fips2 == 47) %>% 
  group_by(year, fouryr_inc_lvl) %>% 
  summarise(avg_comp = mean(fouryr_inc_lvl_size, na.rm = T)) %>% 
  ggplot(aes(x=year, y=avg_comp, color = fouryr_inc_lvl))+
  geom_line()+
  #facet_grid(predomdegawrd~public_private)+
  ggtitle("Put in Shiny:4year cohort average number of students by income level")
```

```{r}

four_yr_piv_gen <- insdta %>% select(id, name, highestdegawrd, predomdegawrd, state_fips2, ownership, public_private, region_id, region, starts_with("Four_yr")) %>% 
  select(id, name, highestdegawrd, predomdegawrd, state_fips2, ownership, public_private, region_id, region, ends_with("male_students")) %>%
  pivot_longer(cols = starts_with("Four"),
               names_to = "fouryr_gen",
               values_to = "fouryr_gen_val")
```

```{r}
four_yr_piv_gen %>% 
  ggplot(aes(x=fouryr_gen, y=fouryr_gen_val, fill=fouryr_gen))+
  geom_bar(position = "dodge", fun = "mean", stat = "summary")+
  facet_grid(predomdegawrd~public_private)
```




```{r}
insdta %>% group_by(year) %>% 
  summarise(comp_fg = mean(Four_yr_completion.first_generation_students, na.rm =T),
            comp_nfg = mean(Four_yr_completion.no_first_generation_students, na.rm = T)) %>% 
  ggplot(aes(x=year))+
  geom_line(aes(y=comp_fg), color = "darkorange")+
  geom_point(aes(y=comp_fg), color = "darkorange")+
  geom_line(aes(y=comp_nfg), color = "darkorchid")+
  geom_point(aes(y=comp_nfg), color = "darkorchid")
```
```{r}
insdta %>% 
  group_by(year) %>% 
  summarize(enrollment = sum(size, na.rm = T)) %>% 
  mutate(Diff_year = year - lag(year),
         Diff_growth = enrollment - lag(enrollment),
         Rate_percent = (Diff_growth/Diff_year)/enrollment * 100)
  
  
```
```{r}
insdta %>% 
  filter(year > 2001) %>% 
  group_by(year, public_private) %>% 
  summarize(enrollment = sum(size, na.rm = T)) %>% 
  
  ggplot()+
  geom_line(aes(x=year, y=enrollment, color = public_private))+
  ylab('Total Enrollment')+
  scale_x_continuous(breaks = seq(2000, 2022, 1))+
  
   theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1),
         legend.title=element_blank())+
  
  ggtitle('Total Enrollment')
  
```





```{r}
enr <- insdta %>% 
  filter(year > 2001) %>% 
  group_by(year) %>% 
  summarize(enrollment = sum(size, na.rm = T)) %>% 
  mutate(Diff_year = year - lag(year),
         Diff_growth = enrollment - lag(enrollment),
         Rate_percent = (Diff_growth/Diff_year)/enrollment) %>% 
  ggplot(aes(x=year, y=Rate_percent))+
  geom_bar(stat = 'identity')+
  ylab('Percent Change')+
  scale_x_continuous(breaks = seq(2000, 2022, 1))+
  scale_y_continuous(labels = percent)+
  theme_minimal()+
   theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1))+
  
  ggtitle('Year to Year Percent Change in Enrollment')

enr
dev.off()

ggsave(enr, width = 10, height = 6, dpi = 300, filename = "enr.pdf")

  
```

```{r}
enr <-
  insdta %>% 
  filter(region_id != 0, region_id != 1, ownership != 3) %>% 
  filter(year > 2001) %>% 
  group_by(public_private, year) %>% 
  summarize(enrollment = sum(size, na.rm = T)) %>% 
  mutate(Diff_year = year - lag(year),
         Diff_growth = enrollment - lag(enrollment),
         Rate_percent = (Diff_growth/Diff_year)/enrollment) %>% 
  ggplot(aes(x=year, y=Rate_percent, fill = public_private))+
  geom_bar(stat = 'identity', position = 'dodge')+
  ylab('Percent Change')+
  scale_x_continuous(breaks = seq(2000, 2022, 1))+
  scale_y_continuous(labels = percent)+
  theme_minimal()+
   theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1),
         legend.title=element_blank())+
  scale_fill_manual(values = c("Private nonprofit" = "yellow3",
                     "Public" = "lightslateblue"))+
  
  ggtitle('Year to Year Percent Change in Enrollment')

enr
dev.off()
# 
ggsave(enr, width = 10, height = 6, dpi = 300, filename = "enrcol.pdf")

  
```

```{r}
enrt <-
  insdta %>% 
  filter(region_id != 0, region_id != 1, ownership != 3) %>% 
  filter(year > 2001) %>% 
  group_by(Predom_Degree_Awarded, year) %>% 
  summarize(enrollment = sum(size, na.rm = T)) %>% 
  mutate(Diff_year = year - lag(year),
         Diff_growth = enrollment - lag(enrollment),
         Rate_percent = (Diff_growth/Diff_year)/enrollment) %>% 
  ggplot(aes(x=year, y=Rate_percent, fill = Predom_Degree_Awarded))+
  geom_bar(stat = 'identity', position = 'dodge')+
  ylab('Percent Change')+
  scale_x_continuous(breaks = seq(2000, 2022, 1))+
  scale_y_continuous(labels = percent)+
  theme_minimal()+
   theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1),
         legend.title=element_blank())+
  scale_fill_manual(values = c("Associate's Degree" = "yellow3",
                     "Bachelor's Degree" = "lightslateblue"))+

  ggtitle('Year to Year Percent Change in Enrollment')

enrt
dev.off()
#
ggsave(enrt, width = 10, height = 6, dpi = 300, filename = "enrdegl.pdf")

  
```



```{r}
percentinc <- insdta %>% 
  filter(region_id != 0, region_id != 9) %>% 
  filter(year >2019) %>% 
  group_by(state, year) %>% 
  summarize(enrollment = sum(size, na.rm = T)) %>% 
  mutate(Diff_year = year - lag(year),
         Diff_growth = enrollment - lag(enrollment),
         Rate_percent = (Diff_growth/Diff_year)/enrollment * 100) %>%
  arrange(desc(Rate_percent)) %>% 
  select(-c(Diff_year, Diff_growth)) %>% head() 
```
```{r}
library(writexl)
```

```{r}
writexl::write_xlsx(percentinc, "percentinc.xlsx")

```


```{r}
percendecc <- insdta %>% 
  filter(region_id != 0, region_id != 9) %>% 
  filter(year >2019) %>% 
  group_by(state, year) %>% 
  summarize(enrollment = sum(size, na.rm = T)) %>% 
  mutate(Diff_year = year - lag(year),
         Diff_growth = enrollment - lag(enrollment),
         Rate_percent = (Diff_growth/Diff_year)/enrollment * 100) %>%
  arrange((Rate_percent)) %>% 
  select(-c(Diff_year, Diff_growth)) %>% head() 
```

```{r}
writexl::write_xlsx(percendecc, "percentdec.xlsx")

```


```{r}
insdta %>% select(predomdegawrd, degrees_awarded.predominant) %>% unique()
```
#of colleges
```{r}
insdta %>% group_by(state) %>% summarise(total = n_distinct(id)) %>% arrange(total)
```
```{r}
library(censusapi)
```



```{r}

state_pop2 <-  getCensus(name="acs/acs5", 
                        vintage=2022,
                        key=census_key, 
                        vars=c("NAME", "B01003_001E"), 
                        region="state:*")
```

```{r}
colnames(state_pop) <- c("state_id", "NAME", "population")
```

```{r}
#write_csv(state_pop, "state_pop.csv")
```


```{r}
inst_data_for_shiny <- read_rds('inst_data3.rds')
```
#join state pop data

```{r}
library(tigris)
```

```{r}
states_b <- states(cb=T)
```

```{r}
library(leaflet)
```

```{r}
states_b %>% 
  leaflet() %>% 
  addTiles %>% 
  addPolygons(popup = ~NAME)
```

#states_merged has population and polygon for state
```{r}
# states_merged <- geo_join(states_b, state_pop, "STATEFP", "state_id")
# states_merged <- subset(states_merged, !is.na(population))
```
```{r}
# pal_sb <- colorNumeric("Greens", domain=states_merged$population)
# 
# 
# leaflet ()%>% 
#   setView(-96, 37.8, 4) %>% 
#   addPolygons(data = states_merged , 
#               fillColor = ~pal_sb(states_merged$population), 
#               fillOpacity = 0.9, 
#               weight = 0.2, 
#               smoothFactor = 0.2
#              ) %>%
#   addLegend(pal = pal_sb, 
#             values = states_merged$population, 
#             position = "bottomright", 
#             title = "Starbucks<br />per 100,000<br/>residents")
```

```{r}
#write_rds(states_merged, 'states_merged.rds')
```

```{r}
#colleges_by_state <- insdta %>% group_by(state_fips2) %>% summarise(n_distinct(id))
```

```{r}
# by_state_summary <- insdta %>% 
#   filter(year ==2021)%>% 
#   group_by(state_fips2) %>% 
#   summarise(no_colleges = n_distinct(id),
#             no_students = sum(size))
# 
# by_state_summary
```
```{r}
by_state_summary_b <- left_join(by_state_summary, state_pop, by = c("state_fips2" = "state_id"))
by_state_summary_b
```

```{r}
#by_state_summary_b <- by_state_summary_b %>% mutate(co_capita = round(no_colleges/population*100000, 2),
                                                    stu_capita = round(no_students/population *100000, 0))
```

```{r}
#write_csv(by_state_summary_b, 'by_state_summary.csv')
```

```{r}
#states_merged_b <-geo_join(states_merged, by_state_summary_b, "STATEFP", "state_fips2")
#states_merged <- subset(states_merged, !is.na(population))
```

```{r}
#write_rds(states_merged_b, 'states_merged_cap.rds')
```

```{r}
#state.name[match(insdta$state, state.abb)]
```

```{r}
prgmsg <- read_csv('prgms.csv')

programlist <- as.list(prgmsg %>% select(starts_with("program_percentage.")) %>% colnames())
programlist <- programlist[-39]
```

```{r}
programlist
```

```{r}
prgmsg %>% select(starts_with("program_percentage.")) %>% rename(gsub("program_percentage.", "")
```

```{r}
library(stringr)
```

```{r}
n <-prgmsg %>% select(starts_with("program_percentage.")) %>% 
  

```

```{r}
prgnames <- read_csv('prgnames.csv')
```

```{r}
prglist <- as.list(prgnames)
```
```{r}
prglist2 <- as.list(prgnames[['name']])
```


```{r}
oldName <- as.list(prgmsg %>% select(starts_with("program_percentage.")) %>% colnames())
```


```{r}
setNames(prgmsg, old = c(oldName),  new =c(prglist2))
```

```{r}
library(data.table)
```

```{r}
xy <- list(prglist2, oldName)
```

```{r}
xy[1, 1]
```

```{r}
prglist2
```

```{r}
library(ggiraph)
```

```{r}

```

